---
title: "HW2"
author: "Lisa He"
date: "4/24/2017"
output: pdf_document
---
1. (a)
```{r 1a}

setwd("/Users/Lisa/Documents/Econ 144/HW 2")
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
z=read.table("labordata.dat")
names(z)= c("male","female","total")
attach(z)
# Convert data to time series format:
male_ts<-ts(male,start=1948,freq=12)
t<-seq(1948, 1991,length=length(male_ts))
female_ts<-ts(female,start=1948,freq=12)
t<-seq(1948, 1991,length=length(female_ts))
total_ts<-ts(total,start=1948,freq=12)
t<-seq(1948, 1991,length=length(total_ts))
#Summary
summary(male_ts)
summary(female_ts)
summary(total_ts)
#Plot
plot(male_ts,ylab="Participation Rate (Male)", xlab="Time", lwd=2, col='skyblue3',ylim=c(75,88), xlim=c(1968,1995))
plot(female_ts,ylab="Participation Rate (Female)", xlab="Time", lwd=2, col='skyblue3',ylim=c(32,58), xlim=c(1968,1995))
plot(total_ts,ylab="Participation Rate (Total)", xlab="Time", lwd=2, col='skyblue3',ylim=c(58,67), xlim=c(1968,1995))

```
(b) Linear:
```{r 1blinear}
#Linear Fit
f1=lm(female_ts~t)
quartz()
par(mfrow=c(2,1))
plot(female_ts,ylab="Participation Rate (Female)", xlab="Time", lwd=2, col='skyblue3',ylim=c(32,58), xlim=c(1968,1995))
#The next commands adds the U.S. recession bands
#nberShade()
lines(t,f1$fit,col="red3",lwd=2)
plot(t,f1$res, ylab="Residuals",type='l',xlab="Time")

```
The linear model is a pretty good fit and the residuals don't look random.

Polynomial:
```{r 1bpoly}
m2=lm(female_ts~t+I(t^2))
quartz()
par(mfrow=c(2,1))
plot(female_ts,ylab="Participation Rate (Female)", xlab="Time", lwd=2, col='skyblue3',ylim=c(32,58), xlim=c(1968,1995))
lines(t,m2$fit,col="red3",lwd=2)
plot(t,m2$res, ylab="Residuals",type='l',xlab="Time")
```
The polynomial model is a pretty good fit and the residuals look random.

Exponential:
```{r 1bexponential}
ds=data.frame(x=t, y=female_ts)
par(mfrow=c(2,1))
plot(female_ts,ylab="Participation Rate (Female)", xlab="Time", lwd=2, col='skyblue3',ylim=c(32,58), xlim=c(1968,1995))
#lines(t,m1$fit,col="green",lwd=2)
m4=nls(y ~ exp(a + b * t),data=ds, start = list(a = 0, b = 0))
lines(ds$x, predict(m4, list(x = ds$x)),col="red3",lwd=2)
plot(t,residuals(m4), ylab="Residuals",type='l',xlab="Time")
summary(m4)
```
The exponentiall model is a pretty good fit and the residuals look random.

(c) linear:
```{r 1clinear}
quartz()
plot(f1$fit,f1$res, pch=20,col="skyblue4",lwd=10)
abline(h=0,lwd=2,col = "red3")
```
The residuals show a pattern where they are positive for the first 10 and last 10 but negative in the middle.

Polynomial:
```{r 1cpoly}
quartz()
plot(m2$fit,m2$res, pch=20,col="skyblue4",lwd=10)
abline(h=0,lwd=2,col = "red3")
```
The residuals show a pattern where they go up and down before 50 and then go down.

Exponential:
```{r 1cexponential}
quartz()
plot(m4)
```
The residuals show a pattern where they go down until 40 and then fluctuate between 40 and 45 and then go up before 50 and then go down.

(d)
```{r 1d}
AIC(f1,m2,m4)
BIC(f1,m2,m4)
```
m2, the polynomial model has the lowest AIC and BIC, so I choose this one.

(e)
```{r 1e}
tn=data.frame(t=seq(1992,2002))
pred=predict(m2, tn, se.fit = TRUE)
#plot(c(male_ts,pred$fit),type='l',xlim=c(1940,2000))

#we are using the prediction function:
pred.plim = predict(m2,tn, level =0.95, interval="prediction")
# simple regression, then use the predict() function
pred.clim = predict(m2, tn,level=0.95, interval="confidence")

#Just overlaying the two prediction so we can compare them
matplot(tn$t,cbind(pred.clim, pred.plim[,-1]),
        lty=c(1,1,1,3,3), type="l", lwd=2, ylab="predicted y",xlab="Time")


```
The point estimate is predicting an upward trend. Both the prediction intervals and the confidence intervals are presenting an upward trend as well.

(f)
```{r 1f}
hwfit<-HoltWinters(female_ts)
quartz()
plot(hwfit)
```
It fits much better than my model.

(g)
```{r 1g}
quartz()
hwpred <- predict(hwfit, 120, prediction.interval = TRUE,level=0.5)
plot(hwfit,hwpred,ylab="Participation Rate (Female)", xlab="Time",xlim=c(1948,2002))
```
It presents an upward trend as well, but the width of the interval increases as the prediction year increases.

2. Problem 3.2
Monthly inflation (growth rate of monthly CPI)
```{r 2a}
rm(list=ls(all=TRUE))
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
require(astsa)
library(xtable)
library(stats)
library(readxl)
library(quantmod)
CPI <- read_excel("~/Documents/Econ 144/HW 2/CPI.xls")
names(CPI)= c("Date","CPI")
attach(CPI)
cpi_ts<-ts(CPI,start=1959,freq=12)
# Growth rates:
log.cpi <- log(cpi_ts)
r_cpi <- diff(log.cpi)
head(log.cpi)
head(r_cpi)
dim(r_cpi)
```
expost real interest rate = 3-month T-bill rate - monthly inflation rate
```{r 2b}
TB3 <- read_excel("~/Documents/Econ 144/HW 2/TB3.xls")
dim(TB3)
names(TB3)= c("Date","TB3")
attach(TB3)
tb3_ts<-ts(TB3,start=1959,freq=12)

#tb3_ts <- tb3_ts[-1:2, ]
dim(tb3_ts)
dim(r_cpi)
realrate<-tb3_ts[,2] - r_cpi[,2]
class(realrate)

```
From HW1
```{r 2c}
# Personal consumption expenditures
PCEC96 <- read_excel("~/Documents/Econ 144/HW 2/Expend.xls")
#dim(PCEC96)
names(PCEC96)= c("Date","Expend")
attach(PCEC96)
pce_ts<-ts(Expend,start=1959,freq=12)
length(pce_ts)
# Growth rates:
log.pce <- log(pce_ts)
r_pce <- diff(log.pce)
length(r_pce)
# Disposable personal income:
DSPIC96 <- read_excel("~/Documents/Econ 144/HW 2/Income.xls")
names(DSPIC96)= c("Date","Income")
attach(DSPIC96)
dpi_ts <- ts(Income, start = 1959, freq = 12)
dpi_ts <- dpi_ts[-698]
length(dpi_ts)
# Growth rates:
log.dpi <- log(dpi_ts)
r_dpi <- diff(log.dpi)

length(r_pce)
length(r_dpi)

realrate_ts <- ts(realrate,start=1959,freq=12)
window(realrate_ts, start = start(realrate_ts) + 1) 
dim(realrate_ts)
length(realrate_ts)
#Regression

#model<-lm(r_pce~r_dpi+realrate, na.action=na.omit)
#summary(model)

```

3. Problem 4.4
```{r 344}
# Clear all variables and prior sessions
rm(list=ls(all=TRUE))
# Load Libraries
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
require(astsa)
library(xtable)
library(stats)
library(readxl)
Table4_1 <- read_excel("~/Documents/Econ 144/HW 2/Problem4_4.xlsx")
names(Table4_1)= c("Date","Price","Rate")
attach(Table4_1)
price_ts<-ts(Price,start=1980,freq=4)
log.price <- log(price_ts)
rate_ts<-ts(Rate,start=1980,freq=4)
log.rate <- log(rate_ts)

```
regression model with one lag 
```{r 3441}
laglogprice <- lag(log.price)
lag1<-lm(log.price ~ laglogprice)
summary(lag1)
```
regression model with two lags 
```{r 3442}
laglogprice2 <- lag(laglogprice)
lag2<-lm(log.price ~ laglogprice+laglogprice2)
summary(lag2)
```
regression model with three lags 
```{r 3443}
laglogprice3 <- lag(laglogprice2)
lag3<-lm(log.price ~ laglogprice+laglogprice2+laglogprice3)
summary(lag3)
```
regression model with four lags 
```{r 3444}
laglogprice4 <- lag(laglogprice3)
lag4<-lm(log.price ~ laglogprice+laglogprice2+laglogprice3+laglogprice4)
summary(lag4)
```


4. Problem 4.6
```{r 446}
# Clear all variables and prior sessions
rm(list=ls(all=TRUE))
# Load Libraries
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
require(astsa)
library(xtable)
library(stats)
library(readxl)
Table4_6 <- read_excel("~/Documents/Econ 144/HW 2/Ex4_6.xlsx")
names(Table4_6)= c("Date","Actual","Forecast")
attach(Table4_6)
actual_ts<-ts(Actual,start=1969,freq=4)
#p<-seq(1, 37,length=length(price_ts))
forecast_ts<-ts(Forecast,start=1969,freq=4)
#r<-seq(1, 37,length=length(rate_ts))
error = actual_ts - forecast_ts
summary(error)
quartz()
ts.plot(actual_ts, forecast_ts, error, col=c(rep("blue",1),rep("black",1),rep("red",1)))
#par(mfrow=c(2,1))
#acf(price_ts,main="Prices")
#acf(diff(price_ts),main="ACF of the First Difference of Prices")
```
Observation: The red line is almost all above 0, which means the actual data is almost always higher than the forecast.

Descriptive Statistics:
```{r 4462}
summary(actual_ts)
summary(forecast_ts)
summary(error)
```
ACF:
```{r 4463}
acf(actual_ts)
acf(forecast_ts)
acf(error)
```
PACF:
```{r 446pacf}
pacf(actual_ts)
pacf(forecast_ts)
pacf(error)
```
Yes, I observe time dependence.

5. Problem 5.4
```{r 554}
# Clear all variables and prior sessions
rm(list=ls(all=TRUE))
# Load Libraries
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
require(astsa)
library(xtable)
library(stats)
library(readxl)
Table4_6 <- read_excel("~/Documents/Econ 144/HW 2/Ex5_4.xlsx")
names(Table4_6)= c("DateA","AMEX","DateS","SP500")
attach(Table4_6)
amex_ts<-ts(AMEX,start=1996,freq=254)
log.amex <- log(amex_ts)
r_amex <- diff(log.amex)
sp500_ts<-ts(SP500,start=1990,freq=239)
log.sp500 <- log(sp500_ts)
r_sp500 <- diff(log.sp500)
par(mfrow=c(2,1))
acf(r_amex,main="AMEX Returns",na.action = na.omit)
acf(r_sp500,main="SP500 Returns",na.action = na.omit)
pacf(r_amex,main="AMEX Returns",na.action = na.omit)
pacf(r_sp500,main="SP500 Returns",na.action = na.omit)

```

They are different from those in Figure 5.6.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

